\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-12-20 NAR_DRAFT.tex
Update significant figures in purity estimate tables.

\begin{table*}[b]
  \tableparts{%
  \caption{Variant calling metrics for all data sets.}
  \label{table:01}%
  }{%
  \begin{tabular}{lllllllllll}
  \toprule
  Scenario & Mutation profile & $\mu$ & \multicolumn{2}{c}{AUROC} & \multicolumn{2}{c}{AUPRC} & \multicolumn{2}{c}{ICI} 
  \\
  & & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
  \\
  \colrule
  100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & .117 & .287
  \\
  100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & .118 & .214
  \\
  100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & .120 & .219
  \\
  500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .138 & .109
  \\
  500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .108 & .112 
  \\
  500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .105 & .116
  \\
  real AML \cite{Griffith2015} & Actual & 3.6e-8 & -- & -- & .995 & .996 & -- & --
  \\
  real breast \cite{Shi2018} & Actual & 7e-7 & -- & -- & .972 & .972 & -- & --
  \\
  \botrule
  \end{tabular}%
  }
  {$\mu$ = per-base mutation rate, AUROC/AUPRC = Area Under Receiver Operating Characteristic / Precision-Recall Curve, ICI = Integrated Calibration Index. Smaller values of ICI are superior.}
 \end{table*}

@@

Shi et al.\ performed multi-region whole exome sequencing on six individual breast tumors to a mean target sequencing depth of 160X and validated all variants identified by three different variant calling pipelines~\cite{Shi2018}.
%We downloaded aligned reads for this tumor, and called variants with MuTect 1.1.7 as described in Materials \& Methods.
We estimated an average per-base mutation rate for these tumor regions of $4\cdot10^{-8}$, which is consistent with observed mutation rates for breast cancers \cite{Alexandrov2019} and with the low number of validated somatic mutations.
For the validated variants, MuTect and \batcave yielded almost identical precision-recall curves (Fig.~\ref{NAR-wes_fig}F and Table~\ref{table:01})

@@

\begin{table*}[b]
  \tableparts{%
  \caption{Sample summary for Shi et al. (2018)}
  \label{table:01}%
  }{%
  \begin{tabular}{lll}
  \toprule
  Sample & Purity estimate (\%) & Mutation rate estimate 
  \\
  \colrule
  Case 1 biorep A & 53.8 & 9.27e-7 
  \\
  Case 1 biorep B & 46.3 & 8.4e-7 
  \\
  Case 1 biorep C & 26.0 & 1.02e-6 
  \\
  Case 2 biorep A & 40.0 & 1.07e-6 
  \\
  Case 2 biorep B & 32.3 & 1.04e-7 
  \\
  Case 2 biorep C & 40(*) & 9.56e-7
  \\
  Case 3 biorep A & 56.4 & 8.7e-7
  \\
  Case 3 biorep B & 72.8 & 7.37e-7 
  \\
  Case 3 biorep C & 64.9 & 4.3e-7 
  \\
  Case 4 biorep A & 60.4 & 8.75e-7 
  \\
  Case 4 biorep B & 61.7 & 7.91e-7 
  \\
  Case 4 biorep C & 57.6 & 8.14e-7 
  \\
  Case 5 biorep A & 78.7 & 5.7e-7 
  \\
  Case 5 biorep B & 76.7 & 5.8e-7 
  \\
  Case 5 biorep C & 80.8 & 1.8e-7 
  \\
  Case 6 biorep A & 47.8 & 2.77e-7 
  \\
  Case 6 biorep B & 49.4 & 4.08e-7 
  \\
  Case 6 biorep C & 52.9 & 4.37e-7 
  \\
  \botrule
  \end{tabular}%
  }
  {Shi et al. did not provide an estimate of purity for Case 2 biorep C. We chose .4 because it is similar to the other biological replicates from this sample.}
\end{table*}

>>>>

\begin{table*}[b]
  \tableparts{%
  \caption{Variant calling metrics for all data sets.}
  \label{table:01}%
  }{%
  \begin{tabular}{lllllllllll}
  \toprule
  Scenario & Mutation profile & $\mu$ & \multicolumn{2}{c}{AUROC} & \multicolumn{2}{c}{AUPRC} & \multicolumn{2}{c}{ICI} 
  \\
  & & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
  \\
  \colrule
  100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & .117 & .287
  \\
  100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & .118 & .214
  \\
  100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & .120 & .219
  \\
  500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .138 & .109
  \\
  500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .108 & .112 
  \\
  500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .105 & .116
  \\
  real AML \cite{Griffith2015} & Actual & 3.6e-8 & -- & -- & .995 & .996 & -- & --
  \\
  real breast \cite{Shi2018} & Actual & 6.6e-7 & -- & -- & .972 & .972 & -- & --
  \\
  \botrule
  \end{tabular}%
  }
  {$\mu$ = per-base mutation rate, AUROC/AUPRC = Area Under Receiver Operating Characteristic / Precision-Recall Curve, ICI = Integrated Calibration Index. Smaller values of ICI are superior.}
 \end{table*}

@@

Shi et al.\ performed multi-region whole exome sequencing on six individual breast tumors to a mean target sequencing depth of 160X and validated all variants identified by three different variant calling pipelines~\cite{Shi2018}.
%We downloaded aligned reads for this tumor, and called variants with MuTect 1.1.7 as described in Materials \& Methods.
\rngcomment{Need to update!}
We estimated an average per-base mutation rate for these tumor regions of $4\cdot10^{-8}$, which is consistent with observed mutation rates for breast cancers \cite{Alexandrov2019} and with the low number of validated somatic mutations.
For the validated variants, MuTect and \batcave yielded almost identical precision-recall curves (Fig.~\ref{NAR-wes_fig}F and Table~\ref{table:01})

@@

\begin{table*}[b]
  \tableparts{%
  \caption{Sample summary for Shi et al. (2018)}
  \label{table:01}%
  }{%
  \begin{tabular}{lll}
  \toprule
  Sample & Purity estimate (\%) & Mutation rate estimate 
  \\
  \colrule
  Case 1 biorep A & 53.8 & 9.3e-7 
  \\
  Case 1 biorep B & 46.3 & 8.4e-7 
  \\
  Case 1 biorep C & 26.0 & 1.0e-6 
  \\
  Case 2 biorep A & 40.0 & 1.1e-6 
  \\
  Case 2 biorep B & 32.3 & 1.0e-7 
  \\
  Case 2 biorep C & 40(*) & 1.0e-6
  \\
  Case 3 biorep A & 56.4 & 8.7e-7
  \\
  Case 3 biorep B & 72.8 & 7.4e-7 
  \\
  Case 3 biorep C & 64.9 & 4.3e-7 
  \\
  Case 4 biorep A & 60.4 & 8.8e-7 
  \\
  Case 4 biorep B & 61.7 & 7.9e-7 
  \\
  Case 4 biorep C & 57.6 & 8.1e-7 
  \\
  Case 5 biorep A & 78.7 & 5.7e-7 
  \\
  Case 5 biorep B & 76.7 & 5.8e-7 
  \\
  Case 5 biorep C & 80.8 & 1.8e-7 
  \\
  Case 6 biorep A & 47.8 & 2.8e-7 
  \\
  Case 6 biorep B & 49.4 & 4.1e-7 
  \\
  Case 6 biorep C & 52.9 & 4.4e-7 
  \\
  \botrule
  \end{tabular}%
  }
  {Shi et al. did not estimate purity for Case 2 biorep C. We chose 40\% purity, based on the other biological replicates from this case.}
\end{table*}
